\chapter{Introduction}
\author{Sebastian Schaffler}

The re-usability of rockets is the future of the aerospace industry, especially in commercial applications such as space tourism and cargo transportation. In order to achieve this goal, traditional spacecraft design decisions have to be rethought to enable mass-production ready solutions that can return from space mostly intact. 

Few commercial aerospace companies have solved these problems to an extent that has allowed them to capitalize on their spacecrafts. While interplanetary tourism is still a couple years away, space transport services are in high demand.

Entering into this market segment is currently only feasible by investing large sums of money on research and development in addition to financial cooperation with an already established space agency.

It is nigh impossible to attempt a market debut from within a non space faring nation and very difficult in those who already have successful space programs. Significantly cheaper solutions are required to prevent this new market from becoming a monopoly.

Solutions and systems developed by public research bodies like academic institutions could help mitigate the upfront cost.

\section{Task}

Launching non-reusable rockets is, apart from safety concerns, considered a largely solved problem. Up until recently landing the boosters of these spacecraft was deemed unfeasible until propulsion based vertical landing became possible through massive advancements in computing capabilities. 

This diploma thesis presents the development of prototype devices able to land and hover with the help of thrusters. The project described in this thesis is intended to be the first of a series of projects with the objective of developing a model rocket capable of vertically launching and landing again. The purpose of this first part is to lay out the foundation for those future projects by building basic testing hardware and developing fundamental back-end software and control algorithms.


%thesis details development of a propulsion based landing system by developing independent prototypes that test the utility of various designs and concepts.

%development of a prototype able to land and hover with the help of thrusters, the project ajdacent to this thesis is intedned to be the first of a series of projects with the end result of developing a model rocket capable of launching and landing again.